---
output:
  html_document: default
  pdf_document: default
---
Prosper Loan Data EDA by MARCO DE FALCO
========================================================


```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(dplyr)
library(gridExtra)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
pl <- read.csv('prosperLoanData.csv')
```

# General Info About the Prosper Loan Data:

Prosper is an online loan platform. The data refers to loans issued from July
2009 onwards. the dataset has many features varying in kind and scope, sometimes
referring to characteristics of the loan itself and some other referring to the
borrowers.

# Univariate Plots Section

To get more general info about the dataset I will run some basic function.
```{r echo=FALSE, Univariate_Plots_1}
dim(pl)
```

81 features seem like a lot of columns!
So the dataset contains almost 114,000 observations across 81 variables. 
For my analysis I will explore only a fraction of these varibles, focusing
solely on those that I found interesting.  More specifically, I will 
investigate the following:  
"BorrowerRate", "ProsperRating..Alpha.", "ProsperScore",
"BorrowerState", "ListingCategory..numeric.", "CurrentDelinquencies", 
"DelinquenciesLast7Years", "StatedMonthlyIncome",
"IncomeRange", "IsBorrowerHomeowner", "TradesNeverDelinquent..percentage.", 
"EmploymentStatus".

```{r echo=FALSE, Univariate_Plots_3}
pl2 <- subset(pl, select=c("BorrowerRate", "ProsperRating..Alpha.", 
                           "ProsperScore",
"BorrowerState", "ListingCategory..numeric.", "CurrentDelinquencies", 
"DelinquenciesLast7Years", "StatedMonthlyIncome",
"IncomeRange", "IsBorrowerHomeowner", "TradesNeverDelinquent..percentage.", 
"EmploymentStatus"))
str(pl2)
```

I have preferred to crete a separate dataframe with the variables of choice.
The 12 variables selected are of different kinds:
Some are categorical: "BorrowerState", "IncomeRange", "IsBorrowerHomeowner"
"ProsperRating..Alpha.".
The rest appears to be intervals, assuming the intervals between values in 
ProsperScore are evenly spaced. In any case, we should
investigate further to gain a better understanding of each.

```{r echo=FALSE, Univariate_Plots_4}
summary(pl2)
```

Above I draw some very basic sumary statistics.
In fact, 
it is useful to have some summary statistics at hand, while we will explore
the different features and correlations.

```{r echo=FALSE, Univariate_Plots_5, message=F, warning=F}
qplot(x = BorrowerRate, data = pl2,  binwidth = .005,color = I("#283149"),
      fill = I("#DA0463")) + xlim(0,quantile(pl2$BorrowerRate, 0.999)) +
      xlab('Borrower Rate')
#half a percent bin
#Plotting 99.5% of the data, excluding the top 0.5%
```

*Borrower Rate* - The plot shows something unexpected: a spike at around 32% 
interest. There are many more loans with a ~32% borrower rate than any other 
rate.
Hopefully we can learn why with further explorations.
I excluded from the plot the top 0.1% rates, as I deemed them outliers.

```{r echo=FALSE, Univariate_Plots_6, message=F, warning=F}
#First we need to change the levels of the Rating
pl2$ProsperRating..Alpha. <- factor(pl2$ProsperRating..Alpha.,
                                    levels=c("AA", "A","B","C", "D","E",
                                             "HR", ""))

#levels(pl2$ProsperRating..Alpha.) <- c("AA", "A","B","C", "D","E",  "HR", "")
qplot(x = ProsperRating..Alpha., data = pl2, color = I("#283149"),
      fill = I("#DA0463")) + xlab('Prosper Rating')
```

*ProsperRating..Alpha.* - The  Prosper Rating assigned at the time the listing
was created: 

N/A, HR, E, D, C, B, A, AA. 

It is important to 
state that the Prosper Rating is 
a different variable from the Prosper Score, yet we don't
know how or why, as they look similar in aim,
so we should further explor their relationship.

```{r echo=FALSE, Univariate_Plots_6b, message=F, warning=F}
qplot(x = as.factor(ProsperScore), data = pl2, color = I("#283149"),
      fill = I("#DA0463")) + xlab('Prosper Score')
```

*ProsperScore* - The description says "ProsperScore is a custom risk score 
built using historical Prosper data. The score ranges from 1-10, with 10 being 
the best, or lowest risk score." This is inconsistent with the plot: Max.:11.00,
and it's no outlier. The Prosper Score has been treated here as a factor.

It will be useful to compare, later in the report *ProsperScore* with
*ProsperRating..Alpha.* as the purpose seems to be the same.


```{r echo=FALSE, Univariate_Plots_7}
BorrowerState <- reorder(pl2$BorrowerState,pl2$BorrowerState,FUN=length)
#sort(table(BorrowerState))
qplot(x = BorrowerState, color = I("#283149"),
      fill = I("#DA0463")) + xlab("Borrower's State") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

*BorrowerState* - It looks like there are twice as many loans from borrowers in 
CA than from any other state. There are many possible
reasons why this has occurred:
Perhaps ProsperLoans is most known
in California, or maybe in some states people tend
to borrow more and so on. Most importantly, California is the most popoulous
state of the US. There are 52 entries, As it includes DC and an unspecified 
value.


```{r echo=FALSE, Univariate_Plots_8}
ListingCategory <- reorder(pl2$ListingCategory..numeric.,pl2$ListingCategory..numeric.,FUN=length)
#sort(table(ListingCategory))
qplot(x = ListingCategory, color = I("#283149"),
      fill = I("#DA0463")) + xlab("Listing Category") + 
  scale_x_discrete(breaks = seq(0, 20, by = 1))
```

*ListingCategory* - The description states:
The category of the listing that the borrower selected when posting their
listing:

0 - Not Available, 1 - Debt Consolidation, 2 - Home Improvement, 3 - Business,
4 - Personal Loan, 5 - Student Use, 6 - Auto, 7- Other,
8 - Baby&Adoption, 9 - Boat, 10 - Cosmetic Procedure, 11 - Engagement Ring,
12 - Green Loans, 13 - Household Expenses, 14 - Large Purchases,
15 - Medical/Dental, 16 - Motorcycle, 17 - RV, 18 - Taxes,
19 - Vacation, 20 - Wedding Loans

I was disappointed from the outcome: I believed I was going to see for
what specific purpose people are borrowing money. 
However, the vast majority asks for a loan to consolidate
(restructure or rearrange), previously issued debt, which tells us nothing about
why they owe money in the first place. Likewise, the second and third highest 
bins read "Not Available" and "Other", far from informative.

I decided to plot the bins in ascending order (rather than in the logical order
0-20) because it's useful to see how certain value counts compare to others.


```{r  echo=FALSE, Univariate_Plots_9, , message=F, warning=F}
CD_l <- qplot(data= pl2, x = log10(pl2$CurrentDelinquencies +1), binwidth = 0.25,
      color = I("#283149"),
      fill = I("#DA0463")) + xlab("Current Delinquencies (Log10 Scale)")

PD_l <-qplot(data= pl2, x = log10(pl2$DelinquenciesLast7Years +1), binwidth = 0.25,
      color = I("#283149"),
      fill = I("#DA0463")) +
  xlab("Delinquencies in the past 7 years (Log10 Scale)") 
grid.arrange(CD_l, PD_l, ncol= 2)
```

*CurrentDelinquencies* - I had to plot this on a Log scale as 
the distribution is
EXTREMELY skewed.
The almost totality of loans are issued to people who do not have current
delinquencies.

*DelinquenciesLast7Years* - A similar skewedness appears here. almost everyone
had
0 delinquencies in their recent history. It is important to say that although
the
rest of the data (the non 0 values) looks centered around 10 in a sort of normal
fashion, in reality we should keep in mind that it's not so "normal" as it's 
plotted on a log scale.

```{r  echo=FALSE, Univariate_Plots_10, , message=F, warning=F}
qplot(data= pl2, x = pl2$TradesNeverDelinquent..percentage., binwidth = 0.01,
      color = I("#283149"),
      fill = I("#DA0463")) +
  xlab("% of never delinquent trades (Log10 Scale)") 
```

*TradesNeverDelinquent..percentage.* - Conversely, if almost everyone has no 
delinquencies (and neither they have had),
the vast majority of loans have been issued
to people that did not trade delinquently. We must rember that a delinquent
trade is a trade performed beyond the delinquency line set for that specific
line of credit. 

```{r  echo=FALSE, Univariate_Plots_11, message=F, warning=F}

qplot(data= pl2, x = pl2$StatedMonthlyIncome, binwidth = 500,
      color = I("#283149"),
      fill = I("#DA0463")) +
  xlim(0,quantile(pl2$StatedMonthlyIncome, 0.99)) +
  xlab("Stated Monthly Income in USD") 
```

*StatedMonthlyIncome* - Self explanatory.
The mode is 3,500 USD. I arbitrarily capped the data at
20,000 USD as there were extreme outliers for borrowers that had 6-digits monthly 
incomes.

```{r  echo=FALSE, Univariate_Plots_12}

pl2$IncomeRange <- factor(pl2$IncomeRange, levels=c("$0", "$1-24,999", "$25,000-49,999",
                                                    "$50,000-74,999", "$75,000-99,999",
                                                    "$100,000+", "Not displayed",
                                                    "Not employed"))
                         
qplot(data= pl2, x = pl2$IncomeRange,
      color = I("#283149"),
      fill = I("#DA0463")) +
  xlab("Borrowers Income Ranges") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

*IncomeRange* - the bulk of the data hovers between 25k USD and 75k USD, which
is actually within the range that I expected.

```{r  echo=FALSE, Univariate_Plots_13}
qplot(data= pl2, x = pl2$EmploymentStatus,
      color = I("#283149"),
      fill = I("#DA0463")) +
  xlab("Borrower Employment Status") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

*EmploymentStatus* - The categories are hard to interpret: It is not clear why
"Employed" should be different from "Full-Time" for instance, rather, it would
be more common sense to consider "Full-Time" as a sub category of Employed.
It is possible that often the data collected from the Borrower was not
 specific enough to indicate the kind of employment.

*IsBorrowerHomeOwner* - I believe it does not make a ton of sense to 
explore this variable visually as the summary table already does a good job in
telling us how the dataset is distributed: 

57478 Borrowers do own a house.

56459 Borrowers do not own a house.

# Univariate Analysis

### What is the structure of your dataset?

The dataset I have subsetted has 113937 listings of loans from the Prosper Loan 
platform. There are 12 features: "BorrowerRate", "ProsperRating..Alpha.", 
"ProsperScore",
"BorrowerState", "ListingCategory..numeric.", "CurrentDelinquencies", 
"DelinquenciesLast7Years", "StatedMonthlyIncome",
"IncomeRange", "IsBorrowerHomeowner", "TradesNeverDelinquent..percentage.", 
"EmploymentStatus".

### What is/are the main feature(s) of interest in your dataset?
The main features of interest in my EDA will be the Borrower Rate together with
the Prosper Rating and the Prosper Score. I want to explore the correlations
between these variables and see how they interact with other features related
to the personal aspects of a borrower's life (more below).

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?
All the other selected variables tell us something about the personal life of
the borrower (e.g. where he lives, his income, why he needs the money, if
he owns a house, if he is a good borrower and a timely payer). I think all
this information might be correlated with the rate at which the borrower pays
back money and the rating/score he has.


### Did you create any new variables from existing variables in the dataset?
No, however, I plan on readjusting the levels of some categorical variables
along the way.

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

As shown, some distributions were extremely skewed. In a few instances I used
a different scal to plot the data in a more meaningful way. In some other 
instances, weird distributions in categorical data were due to bad data or other
issues unknown to the writer.

In some visualizations I have rearranged the order of the entries. 
For IncomeRange
it was necessary as the default visualization provided a random, non progressive
order.

In BorrowerState and ListingCategory..numeric. I adjusted the order to make the 
distribution of data more visible and comparable.

# Bivariate Plots Section

```{r echo=FALSE, Bivariate_Plots , message=F, warning=F}
ggplot(aes(y = pl2$CurrentDelinquencies, x = pl2$DelinquenciesLast7Years),
       data = pl2) + 
 geom_point(alpha = 1/25, color = I('#DA0463'), position = position_jitter())  + 
 coord_trans(y = 'sqrt') +
 geom_line(stat = 'summary', fun.y = mean, color = I("#283149") ) +
  ylim(0, 30) + xlim(0, 70) + xlab("Delinquencies in the Last 7 Years") +
  ylab("Current Delinquencies")

# Limits are arbitrily removing outliers
# DelinquenciesLast7Years caps at 100.
```

One of the things I wanted to first investigate is the correlation between
past and present delinquencies. From the plot above would seem that the more
you have been delinquent in the past, the more you tend to be today. However,
we should keep in mind that we are not performing any statistical test to 
determine the significance of this event.

```{r echo=FALSE, message=F, warning=F}
CD <- ggplot(aes(x = pl2$CurrentDelinquencies, y = pl2$BorrowerRate),
       data = pl2) + xlim(0, 30) +  ylim(0.05, quantile(pl2$BorrowerRate, .995)) +
 geom_point(alpha = 1/25, color = I('#DA0463'), position = position_jitter())  + 
 coord_trans(x = 'sqrt') +
 geom_smooth(color = I("#283149")) + ylab( "Borrower Rate") +
  xlab("Current Delinquencies")
PD <-ggplot(aes(x = pl2$DelinquenciesLast7Years, y = pl2$BorrowerRate),
       data = pl2) + xlim(0, 70) +  ylim(0.05, quantile(pl2$BorrowerRate, .995)) +
 geom_point(alpha = 1/25, color = I('#DA0463'), position = position_jitter())  + 
 coord_trans(x = 'sqrt') +
 geom_smooth(color = I("#283149")) + ylab( "Borrower Rate") +
  xlab("Delinquencies in the Last 7 Years")

grid.arrange(CD,PD,ncol=2)  
```

Does being delinquent affect the rate at which you borrow money? It would look
like there is some positive correlation between your repay rate ande 
both your present and past behaviour in terms of delinquencies. Again, we should
be performing a statistical test in order to reinforce our suspects.

I arbitrarily capped the data above 30 and 70 delinquencies (current and past
respectively), in order to eliminate outliers.

```{r  echo=FALSE, message=F, warning=F}
ggplot(aes(x = as.factor(pl2$ListingCategory..numeric.), y = (pl2$BorrowerRate),
           fill = I('#f3f3f3'), color = I('#DA0463')), 
       data = pl2) +
  geom_boxplot(lwd = 0.50, outlier.color = I("#283149"), outlier.shape = 1) +
  stat_summary(fun.y=mean, colour=I("#283149"), geom="point", 
               shape=18, size=3,show_guide = FALSE) +
  scale_x_discrete() + 
   scale_y_continuous() + 
  xlab('Listing Category') + ylab('Borrower Rate')
```

**An useful legend:**

0 - Not Available, 1 - Debt Consolidation, 2 - 
Home Improvement, 3 - Business,
4 - Personal Loan, 5 - Student Use, 6 - Auto, 7- Other,
8 - Baby&Adoption, 9 - Boat, 10 - Cosmetic Procedure, 11 - Engagement Ring,
12 - Green Loans, 13 - Household Expenses, 14 - Large Purchases,
15 - Medical/Dental, 16 - Motorcycle, 17 - RV, 18 - Taxes,
19 - Vacation, 20 - Wedding Loans

I was stroke by the mean and median rate difference between 9 and 10.
However, before jumping to the conclusion that it's better to take a loan
for a catamaran than for a nose job,  we should take a closer look to the 
*ListingCategory..numeric.* univariate plot above: 9 and 10 are indeed bins
with very few observation, therefore, this high difference in rate can be the
result of chance (e.g. non statistically significant).

Another important thing to notice is that the very frequent 32% rate
is not due to a specific category of loans.

```{r  echo=FALSE}
BRvsPR <- ggplot(aes(x = pl2$ProsperRating..Alpha., y = (pl2$BorrowerRate),
           fill = I('#f3f3f3'), color = I('#DA0463')), 
       data = pl2) +
  geom_boxplot(lwd = 0.50, outlier.color = I("#283149"), outlier.shape = 1) +
   scale_y_continuous() + 
  xlab('Prosper Rating') + ylab('Borrower Rate')

BRvsPS <- ggplot(aes(x = as.factor(pl2$ProsperScore), y = (pl2$BorrowerRate),
           fill = I('#f3f3f3'), color = I('#DA0463')), 
       data = pl2) +
  geom_boxplot(lwd = 0.50, outlier.color = I("#283149"), outlier.shape = 1) +
  scale_x_discrete() + 
   scale_y_continuous() + 
  xlab('Prosper Score') + ylab('Borrower Rate')

grid.arrange(BRvsPR, BRvsPS, ncol = 2)
```

This is interesting. There is a very similar kind of
correlation between the Borrower Rate and either Prosper Rating and Prosper 
Score. This in turn raises another question: If both these variables measure
vitually the same thing, why would they need a double? Maybe they wanted a more
granular view of the same kind of information, using a scale from 1 to 11 
instead of another from HR to AA (7 levels). However, we can't exactly say why,
neither we know why, on average,
a score of 5 is worse than a score of 4 when it comes to 
Borrower Rate.

```{r  echo=FALSE, message=F, warning=F}
ggplot(aes(x = pl2$IsBorrowerHomeowner, y = (pl2$BorrowerRate),
           fill = I('#f3f3f3'), color = I('#DA0463')), 
       data = pl2) +
  geom_boxplot(lwd = 0.50, outlier.color = I("#283149"), outlier.shape = 1) +
  stat_summary(fun.y=mean, colour=I("#283149"), geom="point", 
               shape=18, size=3,show_guide = FALSE) +
  scale_x_discrete() + 
   scale_y_continuous() + 
  xlab('Borrower Home Owner') + ylab('Borrower Rate')
```

As one would expect, it seems that on average, it is better to own
a house when asking for a loan. However, the difference is very slight.
I believe that the ownership of a house is an interesting 
feature and I would like to investigate it more in correlation with other 
variables.

```{r  echo=FALSE, message=F, warning=F}
ggplot(aes(y = (pl2$TradesNeverDelinquent..percentage.), 
           x = as.factor(pl2$ProsperScore),
         fill = I('#f3f3f3'), color = I('#DA0463')), 
       data = pl2) +
  geom_boxplot(lwd = 0.50, outlier.color = I("#283149"), outlier.shape = 1) +
  scale_x_discrete() + 
  stat_summary(fun.y=mean, colour=I("#283149"), geom="point", 
               shape=18, size=3,show_guide = FALSE) +
  coord_flip() +
  xlab('Prosper Score') + ylab('% of never delinquent trades')
```

The higher the Prosper Score, the higher the percentage of positive behavior in
terms of delinquencies. There is an interesting aspect to notice in the plot:
the entries that do not present a Score (labeled NA in the plot) hold the worst 
TradesNeverDelinquent ratio. This fact makes me question the 
idea that the unlabeled data was merely bad data, or entries
that ProsperLoan failed to categorize with a specific score. From this
perspective, "NA" seems like an actual bin, with mean and 
median just a bit below the Score 1.

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

We have discovered some interesting correlations, to summarize a few:

- As delinquencies (present and past) increase, so does the Borrower Rate.

- As Prosper Score and Prosper Rating increase, the Borrower Rate decreases.

- As the proportion of the proportions of non delinquent trades /total trades
approaches 100%, the Prosper Score increases.

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

Yes, amongst the features that lied outside the main variables of interest, I 
found that 
past and current delinquencies are positively correlated with each other.

### What was the strongest relationship you found?

The relation between *BorrowerRate* and *ProsperRating* appears to be the
strongest and the most consistent. Moreover, the symmetry of ProsperScore and
ProsperRating is evident, especially when we relate it to *BorrowerRate*. 
However, we must add that we had not compared the rating and the score with all
the variables of the dataset, therefore, some difference might be there although
not visible up until now.

# Multivariate Plots Section

```{r  echo=FALSE, Multivariate_Plots}
BRvPR_H <- ggplot(aes(x = pl2$ProsperRating..Alpha., y = pl$BorrowerRate,
           color = pl$IsBorrowerHomeowner), data = pl2) + 
 geom_boxplot() + scale_colour_manual(values = c('#DA0463', "#283149")) +
  xlab('Prosper Rating') + ylab('Borrower Rate') +
  labs(color = "Borrower Owns a Home")

BRvIR_H <- ggplot(aes(y = pl2$BorrowerRate, x = pl2$IncomeRange,
                      color = pl$IsBorrowerHomeowner), data = pl2) + 
 geom_boxplot() + scale_colour_manual(values = c('#DA0463', "#283149")) +
  xlab('Income Range') + ylab('Borrower Rate') +
  labs(color = "Borrower Owns a Home") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
grid.arrange(BRvPR_H,BRvIR_H)
```

In the previous section we said that we wanted to investigate more the 
*IsBorrowerHomeOwner* feature together with other variable that appeared to be
highly correlated with BorrowerRate. Above you can see that the difference
(in terms of interest rate) between owning a house or not is almost always
minimal, across all levels of rating. The big difference lies in the not 
labeled data where a strong difference is present.

However, in the plot right below, we see that owning a house does make a 
difference when looking at the interest rate that a borrower is charged, 
against their *IncomeRange*.

```{r  echo=FALSE, message=F, warning=F}
p3 <- ggplot(aes(y = pl2$BorrowerRate, x = pl$StatedMonthlyIncome, color = pl$IsBorrowerHomeowner), data = pl2) + 
  xlim(0, 50000) + 
 geom_point(alpha = 0.02, position = position_jitter()) + 
  scale_colour_manual(values = c('#DA0463', "#283149")) +
  xlab('Stated Monthly Income') + ylab('Borrower Rate') +
  labs(color = "Borrower Owns a Home")

p4 <- ggplot(aes(y = pl2$BorrowerRate, x = pl$StatedMonthlyIncome, color = pl$IsBorrowerHomeowner), data = pl2) + 
  xlim(0, 50000) + 
  geom_density2d() + scale_colour_manual(values = c('#DA0463', "#283149")) +
  xlab('Stated Monthly Income') + ylab('Borrower Rate') +
  labs(color = "Borrower Owns a Home")
grid.arrange(p3,p4)

```

I wanted to substitute the *IncomeRange* variable, with a non-categorical one.
This allows us to draw a different kind of plot,
while garnering more information on
essentially the same question.

The plot above, shows a lot of overplotting, notwithstanding the reduced alpha.
However, we can definitely see that typically the homeowners tend to make more
money than those who do not own a house. The plot right below is extremely
helpful, as we can see that the black and red high density levels, located in
the souther region of the plot, have a non-negligeble difference in terms of
*BorrowerRate*.

```{r  echo=FALSE, message=F, warning=F}
ggplot(aes(y = pl$ProsperScore, x = pl$EmploymentStatus, 
           color = pl$IsBorrowerHomeowner), data = pl2) + 
 geom_boxplot() + scale_colour_manual(values = c('#DA0463', "#283149")) +
xlab('Employment Status') + ylab('Prosper Score') +
  labs(color = "Borrower Owns a Home") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

We said above that PropserScore looks very similar to ProsperRating. They appear
to measure the same thing. However, the advantage of ProsperScore lies in its
numeric nature, meaning that we are not forced to treat it like a factor, which
in turns allows as to easily plot it against a categorical value.

Above we see that within this dataset, owning a house might be harmful 
depending of your employment status. This plot actually raises more questions
than it gives answers: why, for instance, your score would benefit from you not
having a house in case you are jobless or part-time employed?

```{r  echo=FALSE, message=F, warning=F}
ggplot(aes(y = (pl$ProsperRating..Alpha.), x = pl$StatedMonthlyIncome, color = pl$TradesNeverDelinquent..percentage.), data = pl2) + 
  xlim(0, 50000) + 
 geom_point(alpha = .05, position = position_jitter()) +
  scale_y_discrete(limits = c("","HR", "E", "D", "C", "B", "A", "AA")) +
  scale_color_gradient2(high = "#DA0463", midpoint = 0.50, mid = "#404B69",
                       low = "#283149", name ='% of Trades Never Delinquent') +
  xlab('Stated Monthly Income') + ylab('Prosper Rating')
```

I decided to draw this plot primarily to shed more lights on the suspects I had
above when I said that the unlabeled data (in terms of Score) looked like yet 
another bin for very poor performers in terms of *TradesNeverDelinquent*.
As we said before, Rating and Score look very similar in aim, so I decided to
check if this pattern was emerging here as well. As we can see, the more 
blackish
stripe is not HR, but the unlabeled one. This plot also reveals that the highest
rated loans are almost never issued to people without an income (or with a very
low one). Moreover, as one would expect, there are less entries that hold a high
income for the lowest levels of ratings.

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

The most interesting finding in this section, in my opinion is that owning a 
house does not always reduce the borrower interest rate. One would expect that
owning a house would  alwaysdecrease the risk for the lender and 
in turn the interest rate,
but in case you are unemployed, within this dataset
you'd receive an higher rate on average. It is also clear that the variables
*ProsperScore*, *BorrowerRate* and *ProsperRating..Alpha.* are strongly 
correlated with each other: we have said several times that Score and rating
seem to have a similar aim and indeed you see how people not employed have a 
lower score if they own a house.

### Were there any interesting or surprising interactions between features?
The surprising interactions were largely utlined above, however, an extra
finding worth noticing is about that spike in distribution we have noticed
around the 32% interest rate in the univariate section. We could not find an
immediate explanation. It seems clear now that almost all the HR loans hover
around that rate, as if it was the "go-to" interest rate in case you were 
labeled with HR.

------

# Final Plots and Summary

### Plot One

```{r echo=FALSE, Plot_One, message=F, warning=F}

ggplot(aes(y = pl2$BorrowerRate, x = pl$StatedMonthlyIncome, color = pl$IsBorrowerHomeowner), data = pl2) + 
  xlim(0, 50000) + 
  geom_density2d( alpha = .7) + scale_color_manual(values = c('#DA0463', "#283149")) +
  xlab('Stated Monthly Income') + ylab('Borrower Rate') +
  labs(color = "Borrower Owns a Home") + 
  ggtitle('Borrower Rate by Stated Monthly Income by Borrower Home Ownership')
```

### Description One
I decided to focus on this plot  because thanks the density
curves, allow us to identify were the data lumps together:
we see that there are three peaks of distribution around three levels of rates.
These peaks are at around the same levels for both owners and not owners
although the southermost peak highlights a difference between who owns a house
and who does not. 
I believe the curves were an effective way 
to avoid the problems of overplotting.


### Plot Two
```{r echo=FALSE, Plot_Two, message=F, warning=F}
CD <- ggplot(aes(x = pl2$CurrentDelinquencies, y = pl2$BorrowerRate),
       data = pl2) + xlim(0, 30) +  ylim(0.05, quantile(pl2$BorrowerRate, .995)) +
 geom_point(alpha = 1/25, color = I('#DA0463'), position = position_jitter())  + 
 coord_trans(x = 'sqrt') +
 geom_smooth(color = I("#283149")) + ylab( "Borrower Rate") +
  xlab("Current Delinquencies") + 
  ggtitle("Borrower Rate by Current Delinquencies") 
PD <-ggplot(aes(x = pl2$DelinquenciesLast7Years, y = pl2$BorrowerRate),
       data = pl2) + xlim(0, 70) +  ylim(0.05, quantile(pl2$BorrowerRate, .995)) +
 geom_point(alpha = 1/25, color = I('#DA0463'), position = position_jitter())  + 
 coord_trans(x = 'sqrt') +
 geom_smooth(color = I("#283149")) + ylab( "Borrower Rate") +
  xlab("Delinquencies in the Last 7 Years") +
    ggtitle("Borrower Rate by Past Delinquencies") 


grid.arrange(CD,PD,ncol=2) 
```

### Description Two
These two plots were useful to idntify a trend within the dataset: how 
delinquencies correlate with
the interest rate of the borrowers. In general, the higher
the delinquencies (both past and current) the higher the rate, 
although at the same time, at very high levels
of delinquencies the data becomes more and more sparse and the line of tendency
is no longer very clear and accurate. 

### Plot Three
```{r echo=FALSE, Plot_Three, message=F, warning=F}
ggplot(aes(y = (pl$ProsperRating..Alpha.), x = pl$StatedMonthlyIncome, color = pl$TradesNeverDelinquent..percentage.), data = pl2) + 
  xlim(0, 50000) + 
 geom_point(alpha = .05, position = position_jitter()) +
  scale_y_discrete(limits = c("","HR", "E", "D", "C", "B", "A", "AA")) +
  scale_color_gradient2(high = "#DA0463", midpoint = 0.50, mid = "#404B69",
                       low = "#283149", name ='% of Trades Never Delinquent') +
  xlab('Stated Monthly Income') +
  ylab('Prosper Rating')
    ggtitle("Rating by Stated Monthly income by % of Trades Never Delinquent ") 

```

### Description Three

This plot is by far the most interesting in my opinion. It helps us notice that
the Borrower's with the highest ratings very rarely have a 0 USD monthly income.
More importantly, The "purest" shades of red belong to the best ratings. A 
non-banal finding is that unlabeled data ranked as the worst performer
in terms of delinquencies. I believe this to be an added value as it allows us
to shed lights on data that held no label.

------

# Reflection

I started this project with the aim to explore how some of the personal 
characteristics of the borrowers might correlate with the interest rate and
the rating of the borrower. I also wanted to confirm some of the beliefs I had,
which in few cases were disproved (*ex multis* the importance of owning a house)
in bargaining a lower interest rate.

I went through several technical struggles, I had to start ovr several times
as the features I had initially selected did not hold interesting correlations
with each other and ultimately, it took me a lot of trial and error before I 
could find the features that, if explored together, would garner some meaningful
insight.

It was overall great to learn how powerful can R be in drawing compelling
visualizations in so few (sometimes one) line of code.
